---
title: "Class 20150202"
output: html_document
---
iris.melt<-melt(iris, id.var=c("Species"))
ggplot(iris.melt, aes(facotr(variable), value,grou=Species))+geom_boxplot()

```{r}
summary(cars)
```





___

attach(su)
areatable<-table(state,area)
area<-margin.table(areatable,1)
areatable

ddply(su,.(state),numcolwise(sum))
```




ddply(mydata,.(state),numcolwise(sum))
## To calculate, sum of the numbers of nurses and doctors
attach(mydata)
mynurstable<-table(state,nurse)
nurse<-margin.table(mynurstable,1)
nurse
mydocttable<-table(state, doctor)
doctor<-margin.table(mydocttable,1)

```
Then, finally I save the data as a saparate data file. 
```


```

state
---
newdata$zone[newdata$zone == "North-Central"]<-"na"
newdata$zone[newdata$zone == "Northwest"]<-"na"
newdata$zone[newdata$zone == "Northeast"]<-"na"
newdata$zone[newdata$zone == "South-South"]<-"s"
newdata$zone[newdata$zone == "Southeast"]<-"s"
newdata$zone[newdata$zone == "Southwest"]<-"s"

newdata$zone <- as.character(newdata$zone)
str(newdata)
newdata<-na.omit(newdata) ## missing data
finaldata<-subset(mydata, zone == "South-South"|"Southeast"|"Southwest")
finaldata$zone <- as.factor(finaldata$zone)
str(finaldata)
finaldata<-na.omit(finaldata)






You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
